The relative roles of genetic, ecological and environmental factors in species radiations

### 2018-01-16 paper/draft.tex
checkout changes from parallel

%% [NEW ANGLE:]
Island biota have always fascinated and inspired researchers (Darwin, Wallace, ...). %TODO: refs
Nowadays islands are understood as well suitable model systems for ecology and evolution due to their well definable characteristics like isolation. %TODO: ref
However island biota are also increasingly at risk.
While often featuring unique species communities and harboring countless endemics, %TODO: refs
islands are especially threatened by climate change, human impact and invasive species. %TODO: refs
This makes understanding the functioning and evolution of island communities a high priority objective.

Recently, a group of influential island biogeography researchers proposed a set of 50 questions in island biogeography that still need answering (Patiño et al. 2017).
A subset of these questions deals with 
One of the first and to this day highly influential formal theories of island biodiversity is MacArthur and Wilson's seminal equilibrium theory of island biogeography (ETIB). %TODO: ref
In it, species richness on islands is the result of opposing rates of immigration and extinction, both of which in turn are modified by island size (area) and its distance from the mainland (isolation).
This relationship was since successfully applied to a multitude of islands %TODO: refs
, but also island like systems, such as coral reefs, lakes/ponds or forest fragments. %TODO: refs
This general applicability, however, cannot hide the fact that the ETIB fails to deliver a mechanistic explanation of the processes at play.
It has particularly been criticized for its lack of considering speciation. %TODO: refs

>>>>

%% [NEW ANGLE:]
% Island biota have always fascinated and inspired researchers (Darwin, Wallace, ...). %TODO: refs
Islands are understood as suitable model systems for ecology and evolution due to their well definable characteristics like isolation. %TODO: ref
Due to global change, however, island biota are also increasingly at risk.
While often featuring unique species communities and harboring countless endemics, %TODO: refs
islands are especially threatened by climate change, human impact and invasive species. %TODO: refs
This makes understanding the functioning and evolution of island communities a high priority objective.

Recently, a group of influential island biogeography researchers proposed a set of 50 questions in island biogeography that still need answering (Patiño et al. 2017).
A subset of these questions deals with the topic of speciation and radiations on islands.
Still, the proposed mechanisms so far include only ecological or geo-morphological processes,
while genetic properties are hardly recognized in the island context.
At the same time, an increasing number of studies investigates genetic processes and
their effect on reproductive/ecological isolation of populations (e.g., Via).

One of the first and to this day highly influential formal theories of island biodiversity is MacArthur and Wilson's seminal equilibrium theory of island biogeography (ETIB). %TODO: ref
In it, species richness on islands is the result of opposing rates of immigration and extinction, both of which in turn are modified by island size (area) and its distance from the mainland (isolation).
This relationship was since successfully applied to a multitude of islands %TODO: refs
, but also island like systems, such as coral reefs, lakes/ponds or forest fragments. %TODO: refs
This general applicability, however, cannot hide the fact that the ETIB fails to deliver a mechanistic explanation of the processes at play.
It has particularly been criticized for its lack of considering speciation. %TODO: refs
